---
title: "STAT40830 - Assignment 1"
author: "Sabari Karthik S S"
format: beamer-pdf
engine: knitr
execute:
  echo: true
  warning: false
  message: false
fontsize: 10pt
---



```{r}
#Loading the libraries
library(data.table)
library(ggplot2)
```
# Slide 1 – Loading the Datasets

```{r}

fra = fread("indicators_fra.csv")
ind = fread("indicators_ind.csv")
irl = fread("indicators_irl.csv")

dataList = list(France = fra, India = ind, Ireland = irl)
```

For this assignment, I have taken three countries, India( where I am from),
Ireland(where I currently reside) and France (I like Paris)

---

# Slide 2 – Merge and Assign Classes

```{r}
allData = rbindlist(dataList, idcol = "Country")
allData[, Year := as.integer(Year)]
```
Merges all 3 datasets into allData

Converts Year to numeric

---

# Slide 3 – Exploratory Data Analysis

```{r}
# Check dimensions and missing values
cat("Rows:", nrow(allData), "Cols:", ncol(allData), "
")
colSums(is.na(allData))

summary(allData)
str(allData)
```
A basic check of the dataset shows it's well structured:

236,000+ rows across 7 columns covering 3 countries

Only 3 missing values in Year and Value

Year variable ranges from 1960 to 2024

All variable types appear correctly assigned (e.g., character vs numeric)

No major issues detected, so we can safely move into filtering and analysis.
---

# Slide 4 – Filter: GDP per Capita

```{r}
gdp = allData[`Indicator Name` == "GDP per capita (current US$)"]
```

---

# Slide 5 – Plot: GDP per Capita

```{r, fig.width=7, fig.height=4}
library(ggplot2)
ggplot(gdp, aes(Year, `Indicator Value`, color = Country)) +
  geom_line(size = 0.7) +
  labs(title = "GDP per capita (USD)", y = "USD")
```

---

# Slide 6 – Filter: Population Growth

```{r}
pop = allData[`Indicator Name` == "Population growth (annual %)"]
```

---

# Slide 7 – Plot: Population Growth

```{r, fig.width=7, fig.height=4}
ggplot(pop, aes(Year, `Indicator Value`, color = Country)) +
  geom_line(size = 0.7) +
  labs(title = "Population Growth Rate", y = "% growth")
```

---

# Slide 8 – Filter: Education Spend

```{r}
edu = allData[`Indicator Name` == "Government expenditure on education, total (% of GDP)"]
```

---

# Slide 9 – Plot: Education Spend

```{r, fig.width=7, fig.height=4}
ggplot(edu, aes(Year, `Indicator Value`, color = Country)) +
  geom_line(size = 0.7) +
  labs(title = "Education Spend (% GDP)", y = "% of GDP")
```

---

# Slide 10 – Grouped Summary with keyby

```{r, fontsize=\scriptsize}
allData[`Indicator Name` %in% c("GDP per capita (current US$)", "Population growth (annual %)"),
        .(avgValue = mean(`Indicator Value`, na.rm = TRUE)),
        keyby = .(Country, `Indicator Name`)]
```

---

# Slide 11 – Summary of Observations

- Ireland shows highest GDP per capita
- France spends more on education (% of GDP)
- India has the highest population growth rate

---

# Slide 12 – Quarto & Literate Programming

- Report is written entirely in `.qmd`
- All results regenerate with updated data
- Code and narrative co-exist cleanly for clarity

---

# Slide 13 – Version Control Setup

- Project tracked using Git from the start
- Commits mark each step: load, wrangle, plot
- GitHub repo maintained (link available on request)

---

# Slide 14 – Slide Aesthetics & Formatting

- Font size set to 10pt via `.tex` customization
- Used `fig.width`, `fig.height` to scale graphs
- Compact table output using `fontsize=\scriptsize`

---

# Slide 15 – Dependency Management (renv)

```{r eval=FALSE}
# renv::init()
# renv::snapshot()
```

---

# Slide 16 – Final Thoughts

- `data.table` enabled fast, readable code
- Workflow is reproducible and easy to extend
- Framework generalizes well to other indicators/countries

---

# Slide 17 – Thank You!

Any questions?
