---
title: "World Bank Indicators – Assignment 1"
author: "Sabari Karthik S S"
format: beamer-pdf
institute: "UCD – STAT40830"
date: "`r Sys.Date()`"
engine: knitr
execute:
  echo: true
  warning: false
  message: false
fontsize: 10pt
---

# Slide 1 – Load & Merge Datasets

We load and merge three country datasets using `data.table`.

```{r, echo=FALSE}
library(data.table)
library(ggplot2)
fra = fread("indicators_fra.csv")
ind = fread("indicators_ind.csv")
irl = fread("indicators_irl.csv")
dataList = list(France = fra, India = ind, Ireland = irl)
allData = rbindlist(dataList, idcol = "Country")
allData[, Year := as.integer(Year)]
allData[, Value := as.numeric(Value)]
allData = allData[!is.na(Year) & !is.na(Value)]
```

---

# Slide 2 – Data Overview

We summarize structure, missing values, and value ranges.

```{r}
cat("Rows:", nrow(allData), "Cols:", ncol(allData))
colSums(is.na(allData))
summary(allData)
str(allData)
```

---

# Slide 3 – Value Distribution

We visualize distribution of indicator values using histogram.

```{r}
hist(log1p(allData$Value), breaks = 50,
     main = "Histogram of log(1 + Value)",
     xlab = "log1p(Value)")
```

---

# Slide 4 – Most Frequent Indicators

We identify the top 10 most frequently reported indicators.

```{r}
allData[, .N, by = `Indicator Name`][order(-N)][1:10] |>
  ggplot(aes(x = reorder(`Indicator Name`, N), y = N)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  labs(title = "Top 10 Most Common Indicators", x = "Indicator Name", y = "Count") +
  theme_minimal()
```

---

# Slide 5 – GDP per Capita

We compare GDP per capita trends for the three countries.

```{r}
gdp = allData[`Indicator Name` == "GDP per capita (current US$)"]
ggplot(gdp, aes(Year, Value, color = Country)) +
  geom_line(size = 0.7) +
  labs(title = "GDP per capita (USD)", y = "USD") +
  theme(plot.margin = margin(10, 10, 10, 10))
```

Ireland's GDP surged after 2000. France is steady. India is far lower.

---

# Slide 6 – Population Growth

We examine population growth trends across countries.

```{r}
pop = allData[`Indicator Name` == "Population growth (annual %)"]
ggplot(pop, aes(Year, Value, color = Country)) +
  geom_line(size = 0.7) +
  labs(title = "Population Growth Rate", y = "% growth") +
  theme(plot.margin = margin(10, 10, 10, 10))
```

India’s growth declines slowly. France is stable. Ireland fluctuates.

---

# Slide 7 – Education Spend

We track education spending as % of GDP for each country.

```{r}
edu = allData[`Indicator Name` == "Government expenditure on education, total (% of GDP)"]
ggplot(edu, aes(Year, Value, color = Country)) +
  geom_line(size = 0.7) +
  labs(title = "Education Spend (% GDP)", y = "% of GDP") +
  theme(plot.margin = margin(10, 10, 10, 10))
```

France is high and stable. Ireland is more erratic. India is lower but improving.

---

# Slide 8 – Health Spending

We analyze health expenditure trends over time.

```{r}
health = allData[`Indicator Name` == "Current health expenditure (% of GDP)"]
ggplot(health, aes(Year, Value, color = Country)) +
  geom_line(size = 0.7) +
  labs(title = "Health Spending (% GDP)", y = "% of GDP") +
  theme(plot.margin = margin(10, 10, 10, 10))
```

France leads in health spending. Ireland peaks then dips. India remains low.

---

# Slide 9 – Energy Use

We compare energy use per capita between countries.

```{r}
energy = allData[`Indicator Name` == "Energy use (kg of oil equivalent per capita)"]
ggplot(energy, aes(Year, Value, color = Country)) +
  geom_line(size = 0.7) +
  labs(title = "Energy Use per Capita", y = "kg of oil equivalent") +
  theme(plot.margin = margin(10, 10, 10, 10))
```

France and Ireland use more energy per person. India is steadily increasing.

---

# Slide 10 – CO2 Emissions

We analyze CO₂ emissions from the power sector.

```{r}
co2 = allData[`Indicator Name` == "Carbon dioxide (CO2) emissions from Power Industry (Energy) (Mt CO2e)"]
ggplot(co2, aes(Year, Value, color = Country)) +
  geom_line(size = 0.7) +
  labs(title = "CO2 Emissions – Power Industry (Mt CO2e)", y = "Mt CO2e") +
  theme(plot.margin = margin(10, 10, 10, 10))
```

India’s CO₂ emissions from power have grown massively since 1980. France and Ireland remain flat with much lower levels. Gap continues to widen due to India’s rising energy demand.

---

# Slide 11 – Indicator Averages

We summarize average values for key indicators using `keyby`.

```{r, fig.width=6, fig.height=3}
allData[`Indicator Name` %in% c("GDP per capita (current US$)",
                               "Population growth (annual %)",
                               "Government expenditure on education, total (% of GDP)",
                               "Current health expenditure (% of GDP)",
                               "Carbon dioxide (CO2) emissions from Power Industry (Energy) (Mt CO2e)") ,
        .(avgValue = mean(Value, na.rm = TRUE)),
        keyby = .(Country, `Indicator Name`)]
```

---

# Slide 12 – Summary: Country Comparison

We summarize key findings comparing countries overall.

- Ireland leads in GDP per capita, but fluctuates in social spending
- France is consistent across indicators and invests the most in public services
- India has rapid population growth and rising emissions, but low per-capita spending

---

# Slide 13 – Summary: Indicator Perspective

We summarize findings based on individual indicator categories.

- GDP: Ireland highest, India lowest
- Education & Health: France invests the most
- Population: India growing fastest
- Energy Use: France and Ireland higher per capita
- CO₂ Emissions: India’s emissions rising fast from power sector

---

# Slide 14 – Literate Programming

We explain our coding and documentation approach.

- Quarto-based `.qmd` for reproducibility
- Combined code + interpretation
- Easy to update or extend

---

# Slide 15 – Version Control & Styling

We explain tools used to manage the project and format slides.

- Tracked with Git + GitHub
- Font size 10pt, figure widths tuned
- Output formatted clearly

---

# Slide 16 – Dependency Management

We manage packages and versions using `renv`.

```{r eval=FALSE}
# renv::init()
# renv::snapshot()
```

---

# Slide 17 – Final Thoughts

We reflect on the workflow and potential extensions.

- Used `data.table` throughout for speed
- Clean plots using `ggplot2`
- Covered 3 countries and 5 key indicators

---

# Slide 18 – Thank You!
